---
title: "NHL EDA"
subtitle: "Exploratory data analysis using NHL playoff data"  
author: 
  - "Hailey Jensen, Eric Warren, and Gary Zhang"
date: '6/24/2022'
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: "16:9"
      countIncrementalSlides: true
---
```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  #fig.width=9, 
  #fig.height=3.5, 
  fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  #echo = TRUE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE
)
```


```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent(
  base_color = "#7FB0FF"
)
```

# Describing Data

- Our data has 34 columns with 7721 observations
- All observations are from the NHL Stanley Cup 2022 playoffs
  - Not inclusive of the Stanley Cup Finals

```{r, echo = FALSE}
library(tidyverse)
playoff_shot_data <- read_csv("~/R/CMU labs/nhl_playoffs_shots_2022 (1).csv")
playoff_shot_data[1:6, ]%>%
  rename(shooterHand = shooterLeftRight, xcord = arenaAdjustedXCord, ycord = arenaAdjustedYCord)%>%
  select(shooterName, team, shooterHand, event, shotType, shotDistance, teamCode, time, xcord, ycord)%>%
  knitr::kable(format = 'html')
```

---

# Limitations

- Each team played a different number of games
  - For example, Nashville only played 4 games whereas Colorado played 14 games

- We were not given player positions in the data set

- The sample size is small compared to a regular NHL season

- Could not perform in depth player analysis due to only having shot data available 

---

## Background Information

####**What we're doing**

--

Looking at the NHL 2022 playoff data we are presented with a series of questions we wanted to look into more in depth

--

####**Why should you care?**

--

---

class: inverse center middle

# Question 1 

--

Is there a difference between left handed vs. right handed players shots?

---

**The Difference in Shot Location Between Left Handed and Right Handed Shooters**

```{r, echo = FALSE, fig.width= 9, fig.height= 4.9}
playoff_shot_data$shooterLeftRight <- factor(playoff_shot_data$shooterLeftRight, 
                                             labels = c("Left Handed Shooter", "Right Handed Shooter"))

library(sportyR)
playoff_shot_data$shooterLeftRight <- factor(playoff_shot_data$shooterLeftRight, labels = c("Left Handed Shooter", "Right Handed Shooter"))
nhl_shots_filter <- playoff_shot_data %>% filter(arenaAdjustedYCord < 41, arenaAdjustedYCord > -41) %>% mutate(absXCoord = -abs(arenaAdjustedXCord))
geom_hockey(league = "NHL", full_surf = F) + 
  stat_density2d(data = nhl_shots_filter,
                 adjust = 0.5,
                 alpha = 0.5,
                 h = 10,
                 aes(x = absXCoord,
                     y = arenaAdjustedYCord,
                     fill = after_stat(level)),
                 geom = "polygon") +
  scale_fill_gradient(low = "cornflowerblue",
                      high = "darkred") +
  facet_wrap(~ shooterLeftRight, ncol = 2) +
  labs(#title = "Differences in Position on the Ice Both Handed Players Shoot",
       caption = "Data Courtesy of MoneyPuck.com") + 
  theme(legend.position = "right")
```

---
class: inverse center middle

# Question 2 

--

Is there a relationship between shot type and the distance of the shot?

---
**How Type of Shot Varies by Distance**

```{r, echo = FALSE, fig.width= 9, fig.height= 5}
mx <- mean(playoff_shot_data$shotDistance)

playoff_shot_data%>%
  ggplot(aes(x = shotDistance))+
  geom_histogram(bins = 15, 
                 color = "cornflowerblue", 
                 fill = "cornflowerblue",
                 alpha = .22, 
                 size = .65) +
  geom_vline(xintercept= mx,linetype="88", color = "darkred") +
  labs(#title = "How Type of Shot Varies by Distance",
       x = "Distance Away from Goal (in feet)",
       y = "Frequency of Type of Shot",
       caption = "Data Courtesy of Moneypuck.com") +
  theme_bw() +
  facet_wrap(~ shotType, ncol = 1)
```

---
class: inverse center middle

# Question 3

--

idk
---



---
class: center, middle

# Thanks!

Slides created via the R packages:

[**xaringan**](https://github.com/yihui/xaringan)<br>
[gadenbuie/xaringanthemer](https://github.com/gadenbuie/xaringanthemer)

The chakra comes from [remark.js](https://remarkjs.com), [**knitr**](http://yihui.name/knitr), and [R Markdown](https://rmarkdown.rstudio.com).
